### 2020-11-10 scripts/readingdatatrial.Rmd
A start to the BC sandbox

---
title: "Sandbox BC Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r sandbox}
library(here)
library(tidyverse)
library("readxl")

BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))

BCdatastripped <- BCdata %>% 
  select("SCHOOL_YEAR", 
         "DATA_LEVEL", 
         "SPECIAL_NEEDS_CODE", 
         "NUMBER_OF_STUDENTS", 
         "PUBLIC_OR_INDEPENDENT") %>% 
  filter(DATA_LEVEL == "PROVINCE LEVEL", 
         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total")

# view(BCdatastripped)
 ggplot(BCdatastripped, aes(SCHOOL_YEAR, NUMBER_OF_STUDENTS)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
   facet_wrap(~SPECIAL_NEEDS_CODE)

```