### 2020-11-12 scripts/SchoolDistrictCategoriesSandbox_DF.Rmd
Sandbox for School Distric-level Data

---
title: "Sandbox BC Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r sandbox}
library(here)
library(tidyverse)
library("readxl")

BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))

BCdatastripped <- BCdata %>% 
  select("SCHOOL_YEAR", 
         "DATA_LEVEL", 
         "SPECIAL_NEEDS_CODE", 
         "NUMBER_OF_STUDENTS", 
         "PUBLIC_OR_INDEPENDENT",
         "DISTRICT_NUMBER") %>% 
  mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
                          "Physically Dependent" = "A",
                          "Deafblind" = "B",
                          "Moderate+ Intellectual Disability" = "C",
                          "Chronic Health Impairment" = "D",
                          "Visual Impairment" = "E",
                          "Deaf or Hard of Hearing" = "F",
                          "Autism" = "G",
                          "Intensive Behavioural Intervention" = "H",
                          "Mild Intellectual Disability" = "K",
                          "Learning Disability" = "Q",
                          "Gifted" = "P",
                          "Mild Behavioural Intervention" = "R"),
         SCHOOL_YEAR = as.factor(SCHOOL_YEAR),
         NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
  filter(DATA_LEVEL == "DISTRICT LEVEL", 
         DISTRICT_NUMBER == "070",
         #PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
         SPECIAL_NEEDS_CODE != "M",
         SPECIAL_NEEDS_CODE != "J",
         SPECIAL_NEEDS_CODE != "N")

 ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
  geom_line(color = "cornflowerblue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   facet_wrap(~SPECIAL_NEEDS_CODE)

```