### 2020-12-02 scripts/j_play.Rmd

---
title: "j_play"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("papaja")
library(here)
library(tidyverse)
library("readxl")
library(janitor)

#Put library calls you add for your file below here:
```

```{r bc sandbox}
BCdata <- read_excel(here("data", "bcsped_headcount.xlsx"))

BCdatastripped <- BCdata %>% 
  select("SCHOOL_YEAR", 
         "DATA_LEVEL", 
         "SPECIAL_NEEDS_CODE", 
         "NUMBER_OF_STUDENTS", 
         "PUBLIC_OR_INDEPENDENT",
         "DISTRICT_NUMBER") %>% 
  mutate(SPECIAL_NEEDS_CODE = fct_recode(SPECIAL_NEEDS_CODE,
                          "Physically Dependent" = "A",
                          "Deafblind" = "B",
                          "Moderate+ ID" = "C",
                          "Chronic Health Impairment" = "D",
                          "Visual Impairment" = "E",
                          "Deaf or Hard of Hearing" = "F",
                          "Autism" = "G",
                          "Intensive Behavioural Need" = "H",
                          "Mild ID" = "K",
                          "Learning Disability" = "Q",
                          "Gifted" = "P",
                          "Mild Behavioural Need" = "R"),
         SCHOOL_YEAR = as.factor(SCHOOL_YEAR),
         NUMBER_OF_STUDENTS = as.integer(NUMBER_OF_STUDENTS),
         SCHOOL_YEAR = as.numeric(SCHOOL_YEAR)) %>% 
  filter(DATA_LEVEL == "PROVINCE LEVEL",
         PUBLIC_OR_INDEPENDENT == "PROVINCE - Total",
         SCHOOL_YEAR >= 7,
         SPECIAL_NEEDS_CODE != "M",
         SPECIAL_NEEDS_CODE != "J",
         SPECIAL_NEEDS_CODE != "N")

# Making student number column numeric
x <- c(10)
BCdata[ , x] <- apply(BCdata[ , x], 2,
                               function(x)as.numeric(as.character(x)))
#

 ggplot(BCdatastripped, aes(x = SCHOOL_YEAR, y = NUMBER_OF_STUDENTS)) +
  geom_line(color = "cornflowerblue") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
   facet_wrap(~SPECIAL_NEEDS_CODE) +
   labs(title = "BC Provincial Headcount of Students with Exceptionalities by Category",
        x = "School Year",
        y = "Headcount")
```

```{r table: headcount by year}
#Data View
#view(BCdatastripped)
#view(BCdata)
#as_tibble(count_by_year)

#Data Prep
count_by_year <- BCdata %>% 
  janitor::clean_names() %>% 
  group_by(school_year, special_needs_desc) %>% 
  mutate(total_students = sum(number_of_students, na.rm = TRUE)) %>% 
  select("school_year",
         "special_needs_desc",
         "total_students") %>% 
  drop_na()

count_by_year <- BCdata %>% 
  janitor::clean_names() %>% 
  group_by(school_year, special_needs_desc) %>% 
  mutate(total_students = sum(number_of_students, na.rm = TRUE)) %>% 
  select("school_year",
         "special_needs_desc",
         "total_students") %>% 
  pivot_wider(names_from = school_year,
              values_from = total_students) %>% 
  drop_na()

#Table
#apa_table(
#  count_by_year,
#  digits = 0,
#  caption = "Headcount by Classification and District, Select Years 1996-2020",
#  note = "Number of students is estimated and does not include masked data.",
#  col.names = c("School Year",
#                "Special Needs Description",
#                "Number of Students")
#)

#Table is huge - pivot_wider by year and maybe select specific years to look at

#unique(count_by_year$school_year)
```

```{r urban vs rural}
bc_data_district_classification <- read_excel(here("data", "bc_data_district_classification.xlsx"))

district_class <- bc_data_district_classification %>%
  select("SCHOOL_YEAR",
         "DATA_LEVEL",
         "PUBLIC_OR_INDEPENDENT",
         "DISTRICT_NUMBER",
         "DISTRICT_NAME",
         "DISTRICT_POPULATION_2016",
         "DISTRICT_CLASSIFICATION",
         "SPECIAL_NEEDS_CODE",
         "NUMBER_OF_STUDENTS") %>%
  filter(DATA_LEVEL == "DISTRICT LEVEL",
         PUBLIC_OR_INDEPENDENT == "BC Public School") %>%
  pivot_longer(DISTRICT_CLASSIFICATION, values_drop_na = TRUE, values_to = "DISTRICT_CLASSIFICATION") %>% 
    select(-"name")

#Attempting to make columns numeric
view(district_class) #Need cols 6 & 8

x <- c(6,8)
district_class[ , x] <- apply(district_class[ , x], 2,
                               function(x)as.numeric(as.character(x)))
#
```

The district classification data was scaled down to include only public schools, while excluding private institutions. The school districts of Southeast Kootenay, Rocky Mountain, Kootenay Lake, Arrow Lakes, Revelstoke, Kootenay-Columbia, Cariboo-Chilcotin, Sea to Sky, Central Coast, Haida Gwaii, Boundary, Bulkley Valley, Nicola-Simikameen, Peace River South, Peace River North, Gulf Islands, Qualicum, Comox Valley, Campbell River, Gold Trail, Fraser-Cascade, Coast Mountains, Vancouver Island West, Vancouver Island North, Stikine, Nechako Lakes, Nisga'a, and Conseil scolaire francophone were excluded from the data set due to the lack of sufficient population information. The urban vs. rural classifications were made based on the district's population on the 2016 census.  If the population was above 100,000 individuals, it is classified as urban. If the population was below 99,999 individuals, then it is classified as rural.

``` {r District Census Population Table}
#Data Prep
census_data <- district_class %>% 
  janitor::clean_names() %>% 
  group_by(district_name, district_number) %>% 
  mutate(est_students = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c("school_year",
            "data_level",
            "public_or_independent",
            "special_needs_code",
            "number_of_students")) %>% 
  distinct()

#Table
apa_table(
  census_data,
    digits = 0,
  caption = "District Classification by Population According to 2016 Census",
  note = "Number of students does not include masked data.",
  col.names = c("District Number",
            "District Name",
            "Population",
            "Classification",
            "Number of Students")
)
```

```{r rural play}
#Data Prep
rural_districts <- district_class %>% 
  janitor::clean_names() %>% 
  filter(district_classification == "Rural") %>% 
  select(-c("data_level",
            "public_or_independent",
            "district_population_2016"))

#Viz - Totals by Code
ggplot(rural_districts, 
       aes(special_needs_code, number_of_students,
           color = special_needs_code)) +
  geom_line() +
  facet_wrap(~school_year) +
  labs(title = "Total Students by Exceptionality",
       x = "Exceptionality",
       y = "Number of Students")

#Data for Rural Totals (NO categories)
rural_totals <- rural_districts %>% 
  group_by(school_year, district_name, district_number) %>% 
  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
  distinct()

#Data Viz Totals/Time
ggplot(rural_totals, aes(school_year, students)) +
  geom_col(aes(color = district_name)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#What's going on with Prince George? Why is there such a steep drop in 2006?
pg <- rural_districts %>% 
  filter(district_name == "Prince George") %>% 
  as_tibble()

pg_sums <- pg %>% 
  group_by(school_year) %>% 
  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c(district_number, special_needs_code, district_classification, number_of_students)) %>% 
  distinct()

#What's happening by Dx?
pg_dx_sum <- pg %>% 
  group_by(school_year, special_needs_code) %>% 
  mutate(student = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c(district_number, district_classification, number_of_students)) %>% 
  distinct()

ggplot(pg_dx_sum, aes(school_year, student)) +
  geom_col() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~special_needs_code)

#What's going on with gifted?????
pg_gifted <- pg %>% 
  filter(special_needs_code == "P") %>% 
  select(-c(district_number, district_classification))

ggplot(pg_gifted, aes(school_year, number_of_students)) +
  geom_col() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r urban play}
#Data Prep
urban_districts <- district_class %>% 
  janitor::clean_names() %>% 
  filter(district_classification == "Urban") %>% 
  select(-c("data_level",
            "public_or_independent",
            "district_population_2016"))

urban_totals <- urban_districts %>% 
  group_by(school_year, district_name, district_number) %>% 
  mutate(students = sum(number_of_students, na.rm = TRUE)) %>% 
  select(-c(special_needs_code, district_classification, number_of_students)) %>% 
  distinct()

#Data Viz Totals/Time
ggplot(urban_totals, aes(school_year, students)) +
  geom_col(aes(color = district_name)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

#Viz - Totals by Code
ggplot(urban_districts, 
       aes(special_needs_code, number_of_students,
           color = special_needs_code)) +
  geom_line() +
  facet_wrap(~school_year) +
  labs(title = "Total Students by Exceptionality",
       x = "Exceptionality",
       y = "Number of Students")

```

```{r ASD}
asd <- district_class %>% 
  janitor::clean_names() %>% 
  filter(special_needs_code == "G")
```

```{r codes and descs}
code_key <- bc_data_district_classification %>% 
  janitor::clean_names() %>% 
  select(special_needs_code, special_needs_desc) %>% 
  distinct()
```

```{r clean OR data}
primodata <- read_excel(here("data", "primo_merged_oregon_data.xlsx"))
cleanest_primo <-primodata %>% 
  remove_empty(which = c("rows", "cols")) %>% 
  clean_names()

cleanest_primo <- cleanest_primo %>% 
  select(-six_to_twentyone_total, -x24)

cleanest_primo2 <- cleanest_primo [!(cleanest_primo$disability=="TOTAL (SUM OF ALL OF THE ABOVE)" |cleanest_primo$disability=="OTHER HEALTH IMPAIRMENT" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY" | cleanest_primo$disability=="SPECIFIC LEARNING DISABILITY"),]
cleanest_primo2

fresh_primo <- cleanest_primo2 %>% 
  pivot_longer(starts_with("x"), names_to = "age", values_to = "frequency") %>% 
  mutate(age = parse_number(age), 
         year = parse_number(year))

fresh_primo %>% 
  group_by(year, disability) %>% 
  summarize(total=sum(frequency, na.rm=TRUE))

```

```{r Oregon plot}

ggplot(fresh_primo, aes(age, frequency))+
  geom_line(aes(color=disability))+
  facet_wrap(~year)+
    labs(color="disability", x= "Student Age", y="Student Count", title ="Disability Prevalence By Age")
```